### 2019-07-19 manuscript/methods.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the \mbox{CCDBcurator} R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 2n=24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}. %IM suggestion
Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
Following previous work, we coded any species with a functional SI system as $I$, even if SC or dioecy was also reported.
Dioecious species without functional SI were coded as $C$.
% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
%R- Okay- I was getting confused as well thanks for clarifying

@@

In order to investigate the association between ploidy level and diversification, we first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
We therefore define several models that incorporate additional forms of diversification rate heterogeneity.

The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
In this trait-independent model known as CID \citep{beaulieu_2016}, hidden traits can affect diversification but the observed traits do not.
Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.

We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.

We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.

@@

We propose five breeding system models following the same logic as the ploidy models above.
Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.

We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
Similar models were used by \citet{freyman_2019} to study diversification in Onagraceae.

@@

The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.

As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
\citep[similar to][]{caetano_2018, huang_2018}. %IM suggestion
We also fit the analogous models but allowing for diploidization (M21--M25).

@@

In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
To lump states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$.%IM suggestion, describe more the lumping
We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e. $q_0=q_{IC}=rho_I$) and the rates
of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).

@@

Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
In the one-step pathway, the $CP$ state is produced directly from the $ID$ state when whole genome duplication disables SI.
In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16). Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification. %IM suggestion

Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
We insert an artificial division in the $CP$ state, so that one sub-state contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).

First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.

Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
%B: FIXME, we'll hear complaints again about "this," "these," and ... "their." (below)
% R- fixed
We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
 % R- here a reference would be nice  E: Does Jake have a favorite ref? Otherwise, I guess a basic ecology textbook.-DONE
%
% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
% R- couldn't solve to get the exact solutions turns out it is really hard.
%
% E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
%--------------------------------------------------
% \subsubsection{Diploidization as an exploratory hypothesis}
%
% The models described above allow polyploidization but not its reverse.
% We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
% Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
% Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
% Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
% In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
%--------------------------------------------------

@@

Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.- Happens after publication
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

We report posterior distributions for the model parameters in \cref{figure:netdivall}, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations.
% Itay says remove but some reviewers were interested in the distribution of the states. I say leave we don't need to discuss. I'll mention briefly.
Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}. (\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used in the models.
(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
(B) Category groupings into states for each model class.
In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 %B: NOT USING A/B because that confuses panel A and panel B designations unnecessarily. 
Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
}
\label{figure:stateclassifications}
\end{figure}

@@

% E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME
%B: "K=ln(BF(M1,M2))" -> "K=ln(BF($M_{i}$,$M_{ii}$))" because we use M1 and M2 for model names.
\begin{table}[]
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M26. Lumped D/P             & -1463.22                                                          & M26 vs. M16 & 4.10            & M16 (Moderate)                                                       \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M27. Lumped D/P+A/B         & -1417.67                                                          & M27 vs. M23 & 3.678           & M23 (Moderate)                                                       \\
\textbf{M23. ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
M28. Lumped I/C             & -1458.41                                                          & M28 vs. M16 & -0.69           & No evidence                                                          \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M29. Lumped I/C+ A/B        & -1416.60                                                          & M29 vs. M23 & 2.60            & M23 (Moderate)                                                       \\
\textbf{M23. IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for adding a focal trait to a binary state model via Bayes factors. Three-state models are moderately preferred over two-state models. The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
\label{table:lumped}
\end{table}

@@

% E todo: re-run with new rate estimates; add panel letters?
\begin{figure}
    \centering 
    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
        \includegraphics[width=0.9\textwidth]{pathways}
    \caption{
        Contributions of the two pathways to polyploidy.
        The one-step pathway is direct ID$\rightarrow$CP transitions.
        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
        The top panels show the separate contributions of each pathway.
        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
    }
    \label{figure:pathways}
\end{figure}

@@

%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.

\begin{suppfigure}
\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
\caption{Posterior distribution for all the best models with diploidization} % XXX
\label{suppfigure:alldip}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{comparablesDP.pdf}
\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model. Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped} } % XXX
\label{suppfigure:lumpedDP}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{comparablesIC.pdf}
\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model. Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped} } % XXX
\label{suppfigure:lumpedIC}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}} % XXX
\label{suppfigure:asymmetric}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{asrDP.pdf}
\caption{Ancestral state estimation using the maximum a posteriori for each node of theM1. D/P ploidy model} % XXX
\label{suppfigure:DPnodipasr}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
\caption{Posterior distribution for each of the parameters in the M1. D/P model} % XXX
\label{suppfigure:DPnodip}
\end{suppfigure}

>>>>

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the \mbox{CCDBcurator} R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level, but it has 2n=24 chromosomes, and  $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, so we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature  \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012} and original experimental crosses (B.I. and E.E.G., unpub. data). %IM suggestion %B:  I returned the aside about original data. I don't think Itay knows that we performed original crosses (many), and they are incorporated in the data, without being present in "the literature" 
%was originally: Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
Following previous work, we coded any species with a functional SI system as $I$, even if SC or dioecy was also reported.
Dioecious species without functional SI were coded as $C$.
% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
%R- Okay- I was getting confused as well thanks for clarifying

@@

We first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
In this trait-independent model known as CID \citep{beaulieu_2016}, hidden traits can affect diversification but the observed traits do not.
Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.

We fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.

@@

We propose five breeding system models following the same logic as the ploidy models above.
Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.
We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
Similar models were used by \citet{freyman_2019} to study diversification in Onagraceae.

@@

The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.
As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
\citep[similar to][]{caetano_2018, huang_2018}. %IM suggestion
We also fit the analogous models but allowing for diploidization (M21--M25).

@@

In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
Lumping states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
First, we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$. %IM suggestion, describe more the lumping
We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e. $q_0=q_{IC}=rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).

@@

Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
In the one-step pathway, the $CP$ state is produced directly from the $ID$ state when whole genome duplication disables SI.
In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16). 
Our results differ from those of \citet{robertson_2011} in part because our approach relies on a dated phylogeny and models that allow for state-dependent diversification. %IM suggestion

Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
We insert an artificial division in the $CP$ state, so that one sub-state contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
Second, we consider not only transitions between states but also diversification within each state.
State-dependent diversification can change the relative contributions of the two pathways.
For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
% R- couldn't solve to get the exact solutions turns out it is really hard.

@@

Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
Scripts for analyses and key results are available as Supplementary Information. 
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

We report posterior distributions for the model parameters in \cref{figure:netdivall}, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations. 
%R: Itay says remove but some reviewers were interested in the distribution of the states. I say leave we don't need to discuss. I'll mention briefly. %B: I am not sure what this comment refers to (up or down?), but I'm not touching.
Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}. (\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{states.pdf}
\caption{
Character states used in the models.
(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
(B) Category groupings into states for each model class.
In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
}
\label{figure:stateclassifications}
\end{figure}

@@

% E: Could we include a figure of the traits on the tree?  At least as supp info. %FIXME

\begin{table}[]
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M26. Lumped D/P             & -1463.22                                                          & M26 vs. M16 & 4.10            & M16 (Moderate)                                                       \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M27. Lumped D/P+A/B         & -1417.67                                                          & M27 vs. M23 & 3.678           & M23 (Moderate)                                                       \\
\textbf{M23. ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
M28. Lumped I/C             & -1458.41                                                          & M28 vs. M16 & -0.69           & No evidence                                                          \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M29. Lumped I/C+ A/B        & -1416.60                                                          & M29 vs. M23 & 2.60            & M23 (Moderate)                                                       \\
\textbf{M23. IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for adding a focal trait to a binary state model via Bayes factors. Three-state models are moderately preferred over two-state models. The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
\label{table:lumped}
\end{table}

@@

% E: %FIXME: re-run with new rate estimates; add panel letters?
\begin{figure}
    \centering 
    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
        \includegraphics[width=0.9\textwidth]{pathways}
    \caption{
        Contributions of the two pathways to polyploidy.
        The one-step pathway is direct ID$\rightarrow$CP transitions.
        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
        The top panels show the separate contributions of each pathway.
        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
    }
    \label{figure:pathways}
\end{figure}

@@

\begin{suppfigure}
\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
\caption{Posterior distribution for all the best models with diploidization.} % XXX
\label{suppfigure:alldip}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{comparablesDP.pdf}
\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model. Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped}. } % XXX
\label{suppfigure:lumpedDP}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{comparablesIC.pdf}
\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model. Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped}. } % XXX
\label{suppfigure:lumpedIC}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}. } % XXX
\label{suppfigure:asymmetric}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{asrDP.pdf}
\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, D/P ploidy model. } % XXX
\label{suppfigure:DPnodipasr}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
\caption{Posterior distribution for each of the parameters in the M1, D/P ploidy model.} % XXX
\label{suppfigure:DPnodip}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{asrDPAB.pdf}
\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4, D/P+A/B asym model.} % XXX
\label{suppfigure:DPnodipABasr}
\end{suppfigure}

\begin{suppfigure}
\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
\caption{Posterior distribution for each of the parameters in the M4, D/P+A/B asym model.} % XXX
\label{suppfigure:DPnodipAB}
\end{suppfigure}
